---
title: "COMPSCIX 415.2 Homework 1"
author: "Praveen Chander"
date: "June 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#My Github repository for my assignments can be found at this URL: https://github.com/praveenchander/compscix-415-2-assignments.git

```{r results='asis', warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
data("WorldCities")
```
#Q1
Following are the datatypes for the 10 Features from the dataset "WorldCities":

Features: Code, Name, Latitiude, Longitude, Country, CountryRegion, Populaton, RegionCode, Region & Date

Observations: 23,018


```{r warning=FALSE, message=TRUE}
glimpse(WorldCities)
```

Printing the first 200 records
```{r warning=FALSE, message=TRUE}
wc <- head(WorldCities, 200)
print(wc)
```
Printing the Distinct Country Code of the 200 records
```{r warning=FALSE, message=TRUE}
country_col <- WorldCities$country
unique(country_col)
```
#Q2
Printing the Distinct Region values from the entire dataset
```{r warning=FALSE, message=TRUE}
unique(WorldCities$region)
```

#Q3
Printing Unique values using %>% Operator
```{r warning=FALSE, message=TRUE}
WorldCities %>% select(country) %>% unique()
```

#Q4
```{r warning=FALSE, message=TRUE}
WorldCities %>% ggplot(aes(x = population)) +
geom_histogram() +
xlab('City Population')

WorldCities %>% ggplot(aes(x = population)) + geom_histogram() + xlab('Updated Population') + ylab('Updated Count') + ggtitle('Updated Population Details') + theme_bw()

```


