---
title: "COMPSCIX 415.2 Homework 4"
author: "Praveen Chander"
date: "July 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#5.6.7
#Q2
Alternate commands that will give you the same output


```{r warning=FALSE, message=FALSE}
library(dplyr)
library(tidyverse)
library(nycflights13)

not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))
#alternate command for not_cancelled %>% count(dest) 
not_cancelled %>% group_by(dest) %>% summarise(n = sum(!is.na(dest)))

#alternate command for not_cancelled %>% count(tailnum, wt = distance)
not_cancelled %>% group_by(tailnum) %>% tally(wt = distance)
```

#Q4
As the **average departure time increases the chances of the flight being cancelled also increases**, most flights are delayed within 60 minutes

```{r warning=FALSE, message=FALSE}

cancelled_delayed <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(prop_cancelled = mean(cancelled),
            avg_dep_delay = mean(dep_delay, na.rm = TRUE))

ggplot(cancelled_delayed, aes(x = avg_dep_delay, prop_cancelled)) +
  geom_point() +
  geom_smooth()
```

#Q6
When sort is set to TRUE, the result is sorted by Desc (highest to lowest)

```{r warning=FALSE, message=FALSE}
not_cancelled %>% count(dest,sort=T)
```

#10.5
#Q1
A tibble prints to fit the user screen and provides an insight of row numbers in the data set, will also be **explicitly mentioned** mtcars (data frame) is printed differently than flights(tibble) data set

```{r warning=FALSE, message=FALSE}
head(mtcars)
as_tibble(mtcars)
head(flights)
```

#Q2
Compare and contrast between data.frame and Tibble

```{r warning=FALSE, message=FALSE}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

tb <- as_tibble(df)
#Displays NULL as there is no variable with name "x"
tb$x
#Displays same as Data Frame, but in table format
tb[,"xyz"]
#Displays same as Data Frame, but in table format
tb[, c("abc", "xyz")]
```

#Q3
**tbl[[var]]** fetches the column name stored in **"var"**

```{r warning=FALSE, message=FALSE}
tbl <- as.tibble(mtcars)

var <- "mpg"
# Fetches the mpg column values
tbl[[var]]

mtcars %>% print(n_extra = 5)
flights %>% print(n_extra = 19)
dim(flights)

```

#Q6
**print(n_extra = 10)** option controls how many additional column names are printed at the footer of a tibble

```{r warning=FALSE, message=FALSE}
flights %>% print(n_extra = 10)
```
#12.3.3
#Q2
The column names 1999 and 2000 should be **quoted**.

```{r warning=FALSE, message=FALSE}

table4a <- data_frame(
  country = c("Afghanistan", "China"),
  "1999" = c(745, 21222),
  "2000" = c(2666, 31000)
  )

as_tibble(table4a)

# Following code will work when the year is in a quoted text
table4a %>%
  gather('1999', '2000', key = "year", value = "cases")

```

#Q3
A new column with the occurance can be added to differentiate the rows, Hence spread function will be executed.

```{r warning=FALSE, message=FALSE}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
# spread(people, key, value)
# This will throw following error
# Error: Duplicate identifiers for rows (1, 3)
glimpse(people)
people <- tribble(
  ~name,             ~key,    ~value, ~occurance,
  #-----------------|--------|------|------
  "Phillip Woods",   "age",       45, 1,
  "Phillip Woods",   "height",   186, 1,
  "Phillip Woods",   "age",       50, 2,
  "Jessica Cordero", "age",       37, 1,
  "Jessica Cordero", "height",   156, 1
)
spread(people, key, value)
```

#Q4
We have to use **gather** instead of spread.

```{r warning=FALSE, message=FALSE}

preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

gather(preg, sex, count, male, female) %>%
  mutate(pregnant = pregnant == "yes",
         female = sex == "female") 
```

#12.4.3
#Q1
**MERGE** --> Merges the remaining values to the last mapped column  
**FILL** --> Fills missing value "NA" either left or right

```{r warning=FALSE, message=FALSE}

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"),extra="merge")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill="right")

```

#Q2
Need to set it to FALSE if you want to create a new variable, else set it to TRUE as below

```{r warning=FALSE, message=FALSE}

test_dat <- tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill="right")

unite(test_dat, one, two, sep = "", remove=T)
```


#Loading the baby names

```{r warning=FALSE, message=FALSE}

file_path <- 'C:\\Users\\PraveenChander\\Desktop\\pySparking\\Class 3 - Introduction to Data Science - COMPSCIX 415.2\\compscix-415-2-assignments\\Assignment 4\\baby_names.txt'
file_path_csv <- 'C:\\Users\\PraveenChander\\Desktop\\pySparking\\Class 3 - Introduction to Data Science - COMPSCIX 415.2\\compscix-415-2-assignments\\Assignment 4\\baby_names.csv'
txt_baby_names_data <- read.delim(file=file_path, sep ="|", header = TRUE, dec =".")


#GLIMPSE of the data from file
glimpse(txt_baby_names_data)


# Export this file as CSV
csv_data <- read.csv(file=file_path,sep = "|", header = TRUE, dec =".")
write.csv(csv_data,file_path_csv)


# Read the CSV file
csv_data2 <- read_csv(file=file_path_csv)
glimpse(csv_data2)
```
