---
title: "COMPSCIX 415.2 Homework 3"
author: "Praveen Chander"
date: "June 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_package, echo=FALSE}
library('ggplot2')
library('tidyverse')
data(mpg)
data("diamonds")
```
#3.8.1
##Q1
In the Question, The plot is skewed between these 2 continuous variables, Therefore instead of using geom_plot(), if geom_jitter() is used the skeness is reduced and also can be plotted w.r.t a **dimensional variable** like "Displacement" as FACTOR for the color

```{r warning=FALSE, message=TRUE}
ggplot(data=mpg, mapping = aes(x=cty, y=hwy)) +geom_jitter(aes(colour=factor(displ)))

# This can still more be grouped by class of the vehicle
ggplot(data=mpg, mapping = aes(x=cty, y=hwy)) +geom_jitter(aes(colour=factor(displ))) + facet_wrap(~class, nrow = 1, ncol = 7)
```

##Q2
The Parameters that control the amount of jittering are WIDTH & HEIGHT
```{r warning=FALSE, message=TRUE}
ggplot(mpg, aes(cty, hwy)) + geom_jitter(aes(colour=factor(displ)), width = 0.25, height = 0.25)
```

##Q3
--> Unlike geom_jitter(), geom_count() exhibit a **skewed** plot.
--> Count has to be plotted w.r.t any **dimensional** categorical variable.
--> For geom_count() you can plot a graph with *point SIZE propotional with a count of a variable*. This is Best used in conjunction with scale_size_area. Larger counts in a particular portion has a big size point plotted

```{r warning=FALSE, message=TRUE}
ggplot(mpg, aes(cty, hwy)) + geom_jitter(aes(colour=factor(displ)))
ggplot(mpg, aes(cty, hwy)) + geom_count(aes(size = ..prop.., group = displ)) + scale_size_area(max_size = 10)
```

##Q4
Default position adjustment for geom_boxplot() is **position='dodge'**
```{r warning=FALSE, message=TRUE}
X <- ggplot(data = mpg, mapping = aes(x=drv, y=hwy, fill=as.factor(class)))
X + geom_boxplot() # with no position
X + geom_boxplot(position='dodge') # with dodge, looks the same as default
X + geom_boxplot(position='identity')
X + geom_boxplot(position='jitter')
```

#3.9.1
##Q2
labs() when set NULL the labels are not displayed, else will display to the set values for the individual values of **X and Y labels** as below

```{r warning=FALSE, message=TRUE}
# With labs()
bar <- ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = cut),
    show.legend = FALSE,
    width = 1
  ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

# without labs()
bar2 <- ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = cut),
    show.legend = FALSE,
    width = 1
  ) +
  theme(aspect.ratio = 1) + labs(x="test x label", y="test y label")

bar
bar2
```

##Q4
1.What does the plot below tell you about the relationship between city and highway mpg? - *Graph shows at what points the city mileage is worse than highway*
2.Why is coord_fixed() important? - *ensures that one unit on the x-axis is the same length as one unit on the y-axis*
3.What does geom_abline() - *Used to draw the slope*
```{r warning=FALSE, message=TRUE}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed()
```

#4.4
##Q1
my_variable <- 10
my_varıable
Variable name should be given exactly as defined, in this case should be my_var**i**able
```{r warning=FALSE, message=TRUE}
my_variable <- 10
my_variable
```

##Q2
Filtered dataset can be assigned to a variable and that variable can be used in the ggplot
```{r warning=FALSE, message=TRUE}
mpg_8 <- mpg %>% filter(cyl == 8)
ggplot(data = mpg_8) + geom_point(mapping = aes(x = displ, y = hwy))

diamond_above3 <- diamonds %>% filter(carat > 3)
ggplot(data = diamond_above3) + geom_point(mapping = aes(x = carat, y = cut))
```

#5.2.4
##Q1
```{r warning=FALSE, message=TRUE}
library(nycflights13)
#1. Had an arrival delay of two or more hours
filter(flights, arr_delay >= 120)
#2. Flew to Houston (IAH or HOU)
filter(flights, dest=='IAH'|dest=='HOU')
#3. Were operated by United, American, or Delta
filter(flights, carrier %in% c('UA','AA','DL'))
#4. Departed in summer (July, August, and September)
filter(flights, month %in% c(7,8,9))
#5. Arrived more than two hours late, but didn’t leave late
filter(flights, arr_delay>120 & dep_delay==0)
#6. Were delayed by at least an hour, but made up over 30 minutes in flight
filter(flights, dep_delay>=60 & arr_delay<=dep_delay-30)
#7. Departed between midnight and 6am (inclusive)
filter(flights, dep_time>=0000 & dep_time <=0600)
```
##Q3
Flights missing dep_time, all actual flight related variables are also missing data for these records. Therefore, these rows represent the flights that were **CANCELED**
```{r warning=FALSE, message=TRUE}
filter(flights, is.na(dep_time))
```
#5.4.1
##Q1
```{r warning=FALSE, message=TRUE}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
df <- flights %>% dplyr:: select(starts_with("arr"), starts_with("dep"))
df
```
##Q3
If the column name is not available the one_of will throw a warning, but it will still return other variables
```{r warning=FALSE, message=TRUE}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))

```