---
title: "COMPSCIX 415.2 Homework 6"
author: "Praveen Chander"
date: "July 17, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Exercise 1
##Q1, Q2
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(mosaicData)
glimpse(Whickham)

# Q1. Variables in the Dataset
# outcome - survival status after 20 years: a factor with levels Alive Dead
# smoker - smoking status at baseline: a factor with levels No Yes
# age - age (in years) at the time of the first survey

# Q2. Observation and its respresentation
# 1314 observations on women with their smoking status
```

##Q3 - Inference - The person existance(outcome) is directly proportional to smoking status
```{r warning=FALSE, message=FALSE}

# The person existance(outcome) is directly proportional to smoking status

d = Whickham %>% count(smoker, outcome)
ggplot(data = d, mapping = aes(x=smoker, y=n, color=outcome)) + geom_point() + facet_wrap(~outcome, as.table = TRUE)
```

## Q4. Age Group Inference - Most Non-smokers live more than 64 years, whereas smokers die at young age
```{r warning=FALSE, message=FALSE}
# Most Non-smokers live more than 64 years, whereas smokers die at young age

Whickham$new_ages <- case_when(
  Whickham$age <= 44   ~ "Age less Than 44",
  Whickham$age > 44 & Whickham$age <= 64    ~ "Age 44 to 64",
  Whickham$age >  64   ~ "Age greater than 64"
)
dd2 <- Whickham %>% group_by(smoker, outcome, new_ages) %>% summarise(Tot=n())
dd2
ggplot(data = dd2, mapping = aes(x=smoker, y=Tot, color=outcome)) + geom_point() + facet_wrap(~new_ages, as.table = TRUE)

```

#Exercise 2
##Q1, Q2
```{r warning=FALSE, message=FALSE}
# Q1 - Histogram / Density plot for the samples
n <- 10000
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
gamma_samp %>% ggplot() + geom_histogram(aes(x = x), binwidth = 1) + theme_bw()
ggplot(gamma_samp) +
    geom_density(aes(x = x)) +
    geom_rug(aes(x = x, y = 0), position = position_jitter(height = 0))

# Q2 - Mean and Standard Deviation for distribution
mean_samp <- gamma_samp %>% .[['x']] %>% mean()
sd_samp <- gamma_samp %>% .[['x']] %>% sd()
mean_samp
sd_samp
```

## Q3 - Sample size = 30
```{r warning=FALSE, message=FALSE}
n <- 30
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
gamma_samp %>% ggplot() + geom_histogram(aes(x = x), binwidth = 1) + theme_bw()

mean_samp <- gamma_samp %>% .[['x']] %>% mean()
sd_samp <- gamma_samp %>% .[['x']] %>% sd()

mean_samp
sd_samp
```

##Q4 - Sample size 30, repeated for 10000 times
```{r warning=FALSE, message=FALSE}

# create a vector with 10000 NAs
mean_samp <- rep(NA, 10000)
# start a loop
for(i in 1:10000) {
g_samp <- rgamma(30, shape = 1, scale = 2)
mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
mean_samp <- as_tibble(mean_samp)
glimpse(mean_samp)
```

## Q5 - Histogram on mean for 30 Bins (Mean calculated for 10,000 samples)
```{r warning=FALSE, message=FALSE}
mean_samp %>% ggplot() + geom_histogram(aes(x = value), binwidth = 0.5) + theme_bw()
```

## Q6, Q7 - mean and SD for 30 Bins (calculated for 10,000 samples)
```{r warning=FALSE, message=FALSE}
avg_samp <- mean_samp %>% .[['value']] %>% mean()
sd_samp <- mean_samp %>% .[['value']] %>% sd()
avg_samp
sd_samp

#Q7 - The value got from the sampling of 30 Bins for 10000 times, satisfies the Central Limit Theorem
```

##Q8 - Histogram, Mean, SD for 300 Bins (calculated of 10000 samples)
```{r warning=FALSE, message=FALSE}
# create a vector with 10000 NAs
mean_samp <- rep(NA, 10000)
# start a loop
for(i in 1:10000) {
g_samp <- rgamma(300, shape = 1, scale = 2)
mean_samp[i] <- mean(g_samp)
}
# Convert vector to a tibble
mean_samp <- as_tibble(mean_samp)
glimpse(mean_samp)
mean_samp %>% ggplot() + geom_histogram(aes(x = value), binwidth = 0.5) + theme_bw()
avg_samp <- mean_samp %>% .[['value']] %>% mean()
sd_samp <- mean_samp %>% .[['value']] %>% sd()
avg_samp
sd_samp
sd=2/sqrt(300)
sd
# Q8 - The values matches as per the Central Limit Theorem
```
